### 2025-04-29 Phase_II/draft/conclusion.tex
edits

\section{Conclusion}

However, II faces a significant drawback: it captures only the magnitude of the signal via photon correlation, resulting in a loss of phase information. This challenge of phase retrieval in II has been effectively addressed through machine learning techniques, particularly with conditional Generative Adversarial Networks (cGAN). Our study demonstrates that applying cGAN to II data successfully recovers the size, shape, and brightness distribution of a fast-rotating star. Evaluations based on image moments—specifically, the monopole, second, and third-order moments—support the effectiveness of cGAN in achieving accurate image reconstruction from a one-night simulation of II using six baselines. A critical factor in the reconstruction process is the extent of Fourier plane coverage, which depends on the number of available telescopes and the total observing time. With a full night of observation using four telescopes, the brightness distribution could be reconstructed with even greater precision. Future work might explore different observatory layouts to assess their impact on image reconstruction quality, such as integrating the Southern Cherenkov Telescope Array (CTA) with MAGIC.

While the results of this study highlight the significant potential of machine learning—particularly cGAN—for image reconstruction in II, several aspects require further refinement. First, detector efficiencies, which impact the signal-to-noise ratio (SNR) of actual observational data, have not yet been incorporated; addressing these factors will be crucial for more accurate SNR estimation. Second, exploring and comparing alternative methods for image generation could reveal approaches that outperform cGAN in reconstructing stellar images with II. Third, experimenting with different loss functions could provide additional insights into reconstruction quality. Although further testing is needed to refine the GAN and enhance its robustness and reliability, our findings suggest that machine learning is a promising approach for phase reconstruction in II.