### 2025-08-29 Phase_II/draft/response.txt
starting response

|  Reviewer:
|  Dear Editor,
|  Thank you for sending me this manuscript by Rai et al on image
|  reconstruction from intensity interferometry. It addresses an important
|  topic, and may after major revision be suitable for publication in AJ. At a
|  high level, the paper does not demonstrate that the GAN presented actually
|  recovers much of the information about the stars presented; or that it
|  would be able to reconstruct images of other objects or with different uv
|  sampling.
|  Kind regards,
|  The Reviewer
|  
|  Comments
|
|  - Rapidly rotating stars can be fit analytically to these data. It may be
|  worth considering an example of an astrophysical source that would be more
|  interesting to fit, where the data analysis problem is less analytically
|  tractable, such as for complex scenes (spotted stars, or disks, for
|  instance).

Counter-argument: We can justify the choice of gravity-darkened stars.

1. Other image-reconstruction methods have been used with Michelson
   interferometry for this type of source.

2. SII has recently measured ellipticity, so gravity-darkening is a
   natural next step.

|  - On the first page: "However, because the primary observable in 69 II is
|  the electromagnetic field intensity rather than the 70 field amplitude, the
|  phase of the interferometric signal 71 is lost." This is not unique to
|  intensity interferometry - you also only measure the field amplitudes in
|  normal interferometry too. What is different about intensity interferometry
|  is you measure electromagnetic field intensity correlations, which
|  translates to getting fringe visibilities but not fringe phases in the uv
|  plane.

|  - On the second page, there is quite a long discussion of the problem of
|  phase retrieval, including many citations. This is good, though perhaps a
|  little long and could be condensed. But there needs to be a clearer
|  statement of why phase retrieval matters - the Fourier transform is
|  Hermitian, and the consequence is that image modes that have even symmetry
|  under inversion are encoded strictly in amplitudes, and with odd symmetry
|  under inversion in phases. The result is that it intensity interferometry
|  strictly contains no information about odd modes of the image and these
|  must be constrained by the even modes through regularization/priors. This
|  idea needs to be discussed.

|  - I think we are seeing the Hermitian symmetry issue in Figure 13: while
|  the GANs do reasonable reconstructions of the equator-on gravity-darkened
|  stars, in the third and separately in the final examples, you see a pole-on
|  star with significant asymmetry in the ground-truth image, which the GANs
|  completely miss. In this figure, the Difference image is illegible; you
|  should try to unravel this into a residuals plot in 1D by mapping those
|  pixelized residuals back to baselines, and where differences are used in
|  images, use a diverging colourmap like coolwarm or seismic rather than
|  viridis.

|  - Figures should be explained fully in their captions, rather than the
|  layout explained like Figure 13 in Sec 5.1.

|  - This discussion of phase retrieval should be separated in the
|  introduction from the discussion of GANs.

|  - There is quite an extensive lit review on the core ideas of intensity
|  interferometry and their corresponding equations, that could perhaps be
|  condensed.

|  - In Figure 2 we have a single sparse sampling map. It appears to be used
|  in all of the other figures throughout the paper. How well does the network
|  perform with different uv sampling? If it's trained only on one, I would
|  expect very poorly - and it's extremely important to test this and think of
|  ways to extend to a more flexible sampling, because you really don't want
|  to have to train your image reconstruction model separately for every
|  single observation.

Counter-argument: the (u,v) sampling for any given data set is fixed
and known, so if computationally feasible [which it is] you do want to
train on the sampling used.

|  - Figures 4 and 5 could be merged to illustrate ideas better; the linear
|  stretch doesn't really tell us anything. What is the story being told here?
|  I would also suggest the authors consider how their subplots are arranged
|  to make better use of space and avoid small figures embedded in large areas
|  of white space. This comment about spatial arrangement applies to Figures
|  14-16 too. (The labels in those figures don't need "The" in "The Ground
|  Truth").

|  - Figure 6 is too small to be easily legible - could this be a two-column
|  figure, with a clearer colourmap? Anyway - this is likely to need to be
|  explained clearly to an astronomer reader, as it may be confusing why the
|  ground truth and observations are merged. A block diagram is essential,
|  showing as a sequence of operations what the workflow is for training and
|  what the workflow is for inference.

|  - Figures 7-12 occupy a great space with very narrow lines that are hard to
|  read, and do not appear to be especially informative in their current form.
|  The loss functions bounce all over the place and just these trace plots vs
|  epoch don't really tell us anything. If you want to make arguments about
|  stability in training vs hyperparameters, it might be better to extract and
|  parametrize the variability (eg mean, std, skewness moments) and plot this
|  as a function of hyperparameters, rather than this quite confusing set of
|  figures.

|  - Section 4.2 refers to a TensorFlow tutorial, and mentions that
|  modifications are essential - what are the modifications? It doesn't appear
|  to be explained in the subsequent sections. Is any of the paper's code
|  being provided open-source? It is more or less essential for a methods
|  paper mainly about software to provide the code in an appropriate open
|  source format, for reproducibility and extensibility.

|  - gravitational darkening -> gravity darkening

|  - The sentence referring to the "ground truth, which the Discriminator uses
|  to distinguish between real images and those generated by the Generator.
|  During training, the GAN aims to replicate these ground truth images."
|  seems to be confused about GANs in theory. The references to this
|  throughout the paper, including in the context of the merged data, need to
|  be clearer. The discriminator doesn't have access to the ground truth, the
|  discriminator has to distinguish between simulated and ground-truth data.

|  - There appears to be no correlation between the x and y centroids of the
|  star relative to the ground truth. This would be a serious problem? It
|  seems to do ok on the second moments, but no recovery at all of the third
|  moments. This may be related to section 5.3 which seems to lack any account
|  of gravity darkening, which is in the simulations.

Counter-argument: yes the centroids are showing noise -- all values
are around (63.5,63.5) which is the centre of the field -- but
absolute centroid is not a measurable quantity in interferometry

|  - No information is provided in the paper about computer systems. What
|  hardware was used (CPUs? GPUs?), and how long did it take for training and
|  for inference? It is important to quantify computational performance.