### 2025-11-05 Phase_II/draft/FlowDiagram_TikZ.tex
intro_reformatted, II.tex, refs.bib, Paragan.tex, main.tex changed

\documentclass{article}
\usepackage{tikz}
\usepackage[margin=0.7in, left=0.7in]{geometry}
\usetikzlibrary{arrows, arrows.meta, positioning}

\tikzset{
  block/.style = {draw, rectangle, minimum height=1.5em, rounded corners, minimum width=3em, align=center},
  arrow/.style = {->, thick},
  d-arrow/.style={thick, dashed, ->, >=stealth, color=blue}, % dashed arrow style
}

% \tikzset{
%   block/.style={rectangle, draw, text centered, rounded corners, minimum height=1cm},
%   arrow/.style={thick, ->, >=stealth},
%   d-arrow/.style={thick, dashed, ->, >=stealth, color=blue}, % dashed arrow style
% }

\begin{document}
%\title{Flow Diagram of the Conditional GAN}
%\date{}
%\maketitle
%\vspace{-5 cm}
\begin{figure}[htbp]
%\centering
\begin{center}
\begin{tikzpicture}[node distance=0.7cm]

% % Data pipeline

%Nodes
% %\node[block, fill=red!20, anchor=north east]
%   (Lat-vector) at ([xshift= 4cm,yshift=1cm]current bounding box.north east)
%   {Random Noise $z$:\\(Latent Vector)};

% Row-1
\node[block, fill=red!20, minimum width = 3 cm, minimum height = 1 cm] (Lat-vector) {Random Noise $z$ \\(Latent Vector)};

\node[block, fill = yellow!15, right= 1.5 cm of Lat-vector, minimum width=3 cm, minimum height=1.2cm] (gen) {{\large \bf{Generator}} \\$log(G(z|y))$};

% Row-2
\node[block, fill = yellow!10, below = 1cm of gen, minimum width=3 cm, minimum height=1.2cm] (gen-out) {Generated (Fake) Image $\hat x$};

\node[block, fill=red!10, left=of gen-out, minimum width = 3 cm, minimum height = 1 cm] (II-sim) {(Additional Info Labels $y$:\\ Simulated II Observation \\ of the Stellar Object)};

\node[block, fill=blue!15, left=of II-sim, minimum width = 3 cm, minimum height = 1 cm] (gr-truth) {Real Image $x$: Model Images of \\ Stellar Object (here, FRS)};

% Row-3
\node[block, fill=green!20, below = of II-sim, minimum width=5cm] (Input-dataset) {Creation of Input Dataset \\ (Merging of Real Images $x$ OR Generated Images $\hat x$ \\ AND II Simulation Data $y$)};

% Row-4
\node[block, fill=green!20, below= 1 cm of Input-dataset, minimum width=5cm] (split) {Split Dataset \\ (Training, Validation, Testing)};

% Row-5
\node[block, fill=green!15, below left= 1 cm and 0.1 cm of split] (train) {Training Set \\ (Learning Model Parameters)};

\node[block, fill=green!10, below= 1 cm of split] (val) {Validation Set \\ (Tuning Hyperparameters)};

\node[block, fill=green!05, below right= 1 cm and 0.1 cm of split] (test) {Testing Set \\ (For Final Evaluation)};

% Row-6
\node[block, fill = blue!20, below left= 2.5 cm and 1cm of val, minimum width=3 cm, minimum height=1.2cm] (disc) {{\large \bf{Discriminator}} \\ $log (D^*(x|y))$};

% Row-7
\node[block, fill = blue!10, below = 1cm of disc, minimum width=3 cm, minimum height=1.2cm] (disc-out) {Binary Decision \\ Real Image OR Fake Image};

\node[block, fill= yellow!10, below = 10 cm of gen-out, minimum width=5 cm, minimum height=1.2cm] (gen-loss) {{\large{\bf {Generator Loss}}}\\ Update $ log(G(z|y)) \equiv log(1-(D^*(G(z|y)|y)) $};

% Row-8
\node[block, fill= blue!10, below = 1cm of disc-out, minimum width=3 cm, minimum height=1.2cm] (disc-loss) {{\large {\bf {Discriminator Loss}}}\\ Update $log(D^*(x|y))$};

% Row-9
\node[block, fill=teal!20, below= 10 cm of val, minimum width=5cm] (res) {Result / Evaluation \\ (Performance Metrics: Accuracy)};

% % Connections
\draw[arrow] (Lat-vector) -- (gen);
\draw[arrow] (gen) -- (gen-out);
\draw[arrow] (gen-out.south) |- (Input-dataset.east);
\draw[arrow] (gr-truth) -- (II-sim);
\draw[arrow] (gr-truth.south) |- (Input-dataset.west);
\draw[arrow] (II-sim) -- (Input-dataset);
\draw[arrow] (Input-dataset) -- (split);

\draw[arrow] (split.west) -| (train.north);
\draw[arrow] (split.south) -- (val.north);
\draw[arrow] (split.east) -| (test.north);
\path (train.south) ++(0,-0.5) coordinate (p1)
      (p1) ++(0.3,0) coordinate (p2);
\draw[arrow] (train.south) -- (p1) -- (p2) -- (disc.north);
\draw[arrow] (val) |- (disc);
\draw[arrow] (disc) -- (disc-out);
\draw[arrow] (disc-out) -- (disc-loss)

% \path (disc-loss.south) ++(0,-1.0) coordinate (p3)
%       (p3) ++(-3.0,0) coordinate (p4)
%       (p4) ++(0,+8.0) coordinate (p5)
%       (p5) ++(2.5, 0) coordinate (p6)
%       (p6) ++(0,-1.3) coordinate (p7);
% \draw[d-arrow] (disc-loss.south) -- (p3) -- (p4) 
% -- (p5) 
% node[pos=1.0,left=10pt, text=blue, font=\large, rotate=90]{ Discriminator Training \\ / Updating Discriminator Parameters}
% -- (p6) -- (p7);

% \path (disc-loss.south) ++(0,-1.0) coordinate (p3)
%       (p3) ++(-3.0,0) coordinate (p4)
%       (p4) ++(0,+8.0) coordinate (p5)
%       (p5) ++(2.5, 0) coordinate (p6)
%       (p6) ++(0,-1.3) coordinate (p7);

% \draw[d-arrow]
%   (disc-loss.south) -- (p3) -- (p4)
%   -- node[midway, left=15pt, text=blue, font=\large, rotate=90, align=center]
%      {Discriminator Training\\Updating Discriminator Parameters}
%   (p5) -- (p6) -- (p7);

\path (disc-loss.south) ++(0,-1.0) coordinate (p3)
      (p3) ++(-3.0,0) coordinate (p4)
      (p4) ++(0,+8.0) coordinate (p5)
      (p5) ++(2.5, 0) coordinate (p6)
      (p6) ++(0,-1.3) coordinate (p7);

% Draw up to p4 first
\draw[d-arrow]
  (disc-loss.south) -- (p3) -- (p4);

% Now draw the vertical dashed segment with centered label
\draw[d-arrow]
  (p4) -- node[pos=0.9, left=15pt, text=blue, font=\large, rotate=90, align=center]
     {Discriminator Training\\Updating Discriminator Parameters}
  (p5);

% Finish the loop
\draw[d-arrow]
  (p5) -- (p6) -- (p7);

\path (disc-out.east) ++ (2.0,0) coordinate (p8)
      (p8) ++(0,1.0) coordinate (p9);
\draw[arrow] (disc-out.east) -- (p8) -- (p9) -- (gen-loss.west);
\path (gen-loss.south) ++(0,-1.0) coordinate (p10)
      (p10) ++(4.5,0) coordinate (p11)
      (p11) ++(0,15) coordinate (p12);
\draw[d-arrow] (gen-loss.south) -- (p10) -- (p11) 
-- node[midway, right=10pt, text=blue, font = \large, rotate=90, anchor=center] {Generator Training/ updating Generator Parameters}
(p12) -- (gen.east);

\path (test.south) ++(0,-3.4) coordinate (p13)
      (p13) ++(-4.5,0) coordinate (p14);
\draw[arrow] (test.south) -- (p13) -- (p14) -- (res.north);

% \node[block, fill=blue!15, minimum width = 3 cm, minimum height = 1 cm] (gr-truth) {Real Image $x$: Model Images of \\ Stellar Object (here, FRS)};
% \node[block, fill=red!10, right=of gr-truth, minimum width = 3 cm, minimum height = 1 cm] (II-sim) {(Additional Info Labels $y$:\\ Simulated II Observation \\ of the Stellar Object)};
% \node[block, fill=red!20, right=of II-sim, minimum width = 3 cm, minimum height = 1 cm] (Lat-vector) {Random Noise $z$:(Latent Vector)};
% \node[block, fill=green!20, below right = 1 cm and 0.1 cm of gr-truth, minimum width=5cm] (Input-dataset) {Creation of Input Dataset \\ (Merging of Ground Truth Images\\ and II Simulation Data)};
% \node[block, fill=green!20, below=of Input-dataset, minimum width=5cm] (split) {Split Dataset \\ (Training, Validation, Testing)};

% \node[block, fill=green!15, below left=0.6cm and 0.1 cm of split] (train) {Training Set \\ (Learning Model Parameters)};
% \node[block, fill=green!10, below=of split] (val) {Validation Set \\ (Tuning Hyperparameters)};
% \node[block, fill=green!05, below right=0.6cm and 0.5 cm of split] (test) {Testing Set \\ (For Final Evaluation)};

% \node[block, fill = blue!15, below left= 3 cm and 0.1 cm of val, minimum width=5 cm, minimum height=1.2cm] (disc) {Discriminator \\ $log (D^*(x|y))$};

% \node[block, fill = blue!10, below = 1cm of disc, minimum width=5 cm, minimum height=1.2cm] (disc-out) {Binary Decision \\ Real Image OR Fake Image};

% \node[block, fill= blue!10, below = 1cm of disc-out, minimum width=5 cm, minimum height=1.2cm] (disc-loss) {Discriminator Loss\\ Update $log(D^*(x|y))$};

% \node[block, fill = yellow!15, below right = 1 cm and 0.5 cm of val, minimum width=5 cm, minimum height=1.2cm] (gen) {Generator \\$G(z|y) \equiv log(1-(D^*(G(z|y)|y))$};

% \node[block, fill = yellow!10, below = 1cm of gen, minimum width=5 cm, minimum height=1.2cm] (gen-out) {Generated (Fake) Image};

% \node[block, fill= yellow!10, below = 1cm of gen-out, minimum width=5 cm, minimum height=1.2cm] (gen-loss) {Generator Loss\\ Update $G(z|y)$};

% % Final result block
% \node[block, fill=teal!20, below= 10 cm of val, minimum width=5cm] (res) {Result / Evaluation \\ (Performance Metrics: Accuracy)};

% % Connections
% \draw[arrow] (gr-truth) -- (II-sim);
% \draw[arrow] (gr-truth.south) |- (Input-dataset.west);
% \draw[arrow] (II-sim) -- (Input-dataset);
% \draw[arrow] (Input-dataset) -- (split); 
% \draw[arrow] (Input-dataset) -- (split);

% \draw[arrow] (split.west) -| (train.north);
% \draw[arrow] (split.south) -- (val.north);
% \draw[arrow] (split.east) -| (test.north);
% \path (train.south) ++(0,-0.5) coordinate (p1)
%       (p1) ++(0.2,0) coordinate (p2);
% \draw[arrow] (train.south) -- (p1) -- (p2) -- (disc.north);
% \draw[arrow] (val) |- (disc);
% \draw[arrow] (disc) -- (disc-out);
% \draw[arrow] (disc-out) -- (disc-loss)

% \path (disc-loss.south) ++(0,-1.0) coordinate (p3)
%       (p3) ++(-3.0,0) coordinate (p4)
%       (p4) ++(0,+8.0) coordinate (p5)
%       (p5) ++(2.5, 0) coordinate (p6)
%       (p6) ++(0,-1.3) coordinate (p7);
% \draw[d-arrow] (disc-loss.south) -- (p3) -- (p4) 
% -- (p5) 
% node[pos=1.0,left=7pt, text=blue, font=\large, rotate=90]{Iteration / Updating Discriminator Parameters}
% -- (p6) -- (p7);

% \path(II-sim.south) ++ (0,-0.5) coordinate (p8)
%      (p8) ++ (7.0,0) coordinate (p9)
%      (p9) ++ (0, -5.5) coordinate (p10)
%      (p10) ++ (-1.5,0) coordinate (p11)
%      (p11) ++ (0, -0.6) coordinate (p12);
% \draw[arrow] (II-sim.south) -- (p8) -- (p9) -- (p10) -- (p11) -- (p12);

% \path(Lat-vector.south) ++ (0, -0.4) coordinate (p13)
%      (p13) ++ (2.0,0) coordinate (p14)
%      (p14) ++ (0,-5.9) coordinate (p15)
%      (p15) ++ (-1.5,0) coordinate (p16)
%      (p16) ++ (0,-0.5) coordinate (p17);
% \draw[arrow] (Lat-vector.south) -- (p13) -- (p14) -- (p15) -- (p16) -- (p17);
% \draw[arrow] (gen) -- (gen-out);
% \draw[arrow] (gen-out) -- (gen-loss);
% \path (gen-out.west) ++ (-1.0,0) coordinate (p18)
%       (p18) ++ (0,0.1) coordinate (p19)
%       (p19) ++ (-4.1,0) coordinate (p20);
% \draw[arrow] (gen-out.west) -- (p18) -- (p19) -- (p20);
%\draw[arrow] (gen-out.west) -- (disc);

\end{tikzpicture}
\end{center}
\end{figure}

\end{document}