### 2025-11-12 Phase_II/draft/response.txt
edits

|  Reviewer:
|  Dear Editor,
|  Thank you for sending me this manuscript by Rai et al on image
|  reconstruction from intensity interferometry. It addresses an important
|  topic, and may after major revision be suitable for publication in AJ. At a
|  high level, the paper does not demonstrate that the GAN presented actually
|  recovers much of the information about the stars presented; or that it
|  would be able to reconstruct images of other objects or with different uv
|  sampling.
|  Kind regards,
|  The Reviewer
|  
|  Comments
|
|  - Rapidly rotating stars can be fit analytically to these data. It may be
|  worth considering an example of an astrophysical source that would be more
|  interesting to fit, where the data analysis problem is less analytically
|  tractable, such as for complex scenes (spotted stars, or disks, for
|  instance).

In this first attempt, is seems reasonable to consider only
gravity-darkened stars, for the following reasons.

1. Other image-reconstruction methods have been used with Michelson
   interferometry for this type of source.

The response to this in the Introduction:

Image reconstruction in gravity darkened fast rotating stars has long been examined using various methods in MI \citep{vanBelle2001, DomicianodeSouza2003, DomicianodeSouza2005, Monnier2007, Pedretti2009, Zhao2009, Martinez2021}.

2. SII has recently measured ellipticity, so gravity-darkening is a
   natural next step.

The response to this in the Introduction:

Recently photosphere oblateness of $\gamma$-Cassiopeia \citep{2025arXiv250615027A} has been measured at the VERITAS observatory using II. These results, and especially, that by Archer et al. \cite{2025arXiv250615027A}  put our work in context. Thus our simulation is a natural next step of the work of Archer et al and others.

|  - On the first page: "However, because the primary observable in 69 II is
|  the electromagnetic field intensity rather than the 70 field amplitude, the
|  phase of the interferometric signal 71 is lost." This is not unique to
|  intensity interferometry - you also only measure the field amplitudes in
|  normal interferometry too. What is different about intensity interferometry
|  is you measure electromagnetic field intensity correlations, which
|  translates to getting fringe visibilities but not fringe phases in the uv
|  plane.

Clarify that we mean phase of the complex visibility. However, we have rephrased the line
to be more clear.

|  - On the second page, there is quite a long discussion of the problem of
|  phase retrieval, including many citations. This is good, though perhaps a
|  little long and could be condensed. But there needs to be a clearer
|  statement of why phase retrieval matters - the Fourier transform is
|  Hermitian, and the consequence is that image modes that have even symmetry
|  under inversion are encoded strictly in amplitudes, and with odd symmetry
|  under inversion in phases. The result is that it intensity interferometry
|  strictly contains no information about odd modes of the image and these
|  must be constrained by the even modes through regularization/priors. This
|  idea needs to be discussed.

Text has been revised.

|  - I think we are seeing the Hermitian symmetry issue in Figure 13: while
|  the GANs do reasonable reconstructions of the equator-on gravity-darkened
|  stars, in the third and separately in the final examples, you see a pole-on
|  star with significant asymmetry in the ground-truth image, which the GANs
|  completely miss. In this figure, the Difference image is illegible; you
|  should try to unravel this into a residuals plot in 1D by mapping those
|  pixelized residuals back to baselines, and where differences are used in
|  images, use a diverging colourmap like coolwarm or seismic rather than
|  viridis.

We increased the number of training, validation and testing dataset. 
Also we increase the training time from 60k to 100k and result has 
been improvised. The difference plot is now in 1-D.

|  - Figures should be explained fully in their captions, rather than the
|  layout explained like Figure 13 in Sec 5.1.

Each figure's captions have been revised now.

|  - This discussion of phase retrieval should be separated in the
|  introduction from the discussion of GANs.

It has been separated now.

|  - There is quite an extensive lit review on the core ideas of intensity
|  interferometry and their corresponding equations, that could perhaps be
|  condensed.

The equations have been condensed now.

|  - In Figure 2 we have a single sparse sampling map. It appears to be used
|  in all of the other figures throughout the paper. How well does the network
|  perform with different uv sampling? If it's trained only on one, I would
|  expect very poorly - and it's extremely important to test this and think of
|  ways to extend to a more flexible sampling, because you really don't want
|  to have to train your image reconstruction model separately for every
|  single observation.

The (u,v) sampling for any given data set is fixed and known.
Training the network for a single observing run is a tolerable
computational cost.  Hence we train on the sampling used.

|  - Figures 4 and 5 could be merged to illustrate ideas better; the linear
|  stretch doesn't really tell us anything. What is the story being told here?
|  I would also suggest the authors consider how their subplots are arranged
|  to make better use of space and avoid small figures embedded in large areas
|  of white space. This comment about spatial arrangement applies to Figures
|  14-16 too. (The labels in those figures don't need "The" in "The Ground
|  Truth").

Both figs 4 and 5 is merged and only linear scale is there. Now the visibility 
pattern is visible as we choose smaller source shown in fig~3. 
The labels in fig 14-16 have been modified now.

|  - Figure 6 is too small to be easily legible - could this be a two-column
|  figure, with a clearer colourmap? Anyway - this is likely to need to be
|  explained clearly to an astronomer reader, as it may be confusing why the
|  ground truth and observations are merged. A block diagram is essential,
|  showing as a sequence of operations what the workflow is for training and
|  what the workflow is for inference.

Figure 6 is bigger now and have been explained that why it is merge.
There is also a block diagram created which explain the image reconstruction using GAN.

|  - Figures 7-12 occupy a great space with very narrow lines that are hard to
|  read, and do not appear to be especially informative in their current form.
|  The loss functions bounce all over the place and just these trace plots vs
|  epoch don't really tell us anything. If you want to make arguments about
|  stability in training vs hyperparameters, it might be better to extract and
|  parametrize the variability (eg mean, std, skewness moments) and plot this
|  as a function of hyperparameters, rather than this quite confusing set of
|  figures.

Figures and text accordingly have been revised.

|  - Section 4.2 refers to a TensorFlow tutorial, and mentions that
|  modifications are essential - what are the modifications? It doesn't appear
|  to be explained in the subsequent sections. Is any of the paper's code
|  being provided open-source? It is more or less essential for a methods
|  paper mainly about software to provide the code in an appropriate open
|  source format, for reproducibility and extensibility.

Give github repo, maybe give most recent version a DOI.

|  - gravitational darkening -> gravity darkening

Modified

|  - The sentence referring to the "ground truth, which the Discriminator uses
|  to distinguish between real images and those generated by the Generator.
|  During training, the GAN aims to replicate these ground truth images."
|  seems to be confused about GANs in theory. The references to this
|  throughout the paper, including in the context of the merged data, need to
|  be clearer. The discriminator doesn't have access to the ground truth, the
|  discriminator has to distinguish between simulated and ground-truth data.

It has been rephrased and clear now. Yes, the Discriminator doesn't have access to the ground truth.
Even Discriminator loss function use Ground truth and generated images to give feedback to the Discriminator

|  - There appears to be no correlation between the x and y centroids of the
|  star relative to the ground truth. This would be a serious problem? It
|  seems to do ok on the second moments, but no recovery at all of the third
|  moments. This may be related to section 5.3 which seems to lack any account
|  of gravity darkening, which is in the simulations.

Counter-argument: yes the centroids are showing noise -- all values
are around (63.5,63.5) which is the centre of the field -- but it is
not important, because absolute centroid is not a measurable quantity
in interferometry.

|  - No information is provided in the paper about computer systems. What
|  hardware was used (CPUs? GPUs?), and how long did it take for training and
|  for inference? It is important to quantify computational performance.

The information is provided and we use CPU for training. This work has used 2 nodes with 96 threads each,
which takes around 20 hours to train the model with 100000 training steps.