### 2025-11-27 Phase_II/draft/Result.tex

\section{\texorpdfstring{\textcolor{blue}{Image Reconstruction: Results and Evaluation}}
                      {Image Reconstruction: Results and Evaluation}}

%section{\textcolor{blue}{Image Reconstruction: Results and Evaluation}}
%\section{\st{Images Reconstructed by the GAN}\textcolor{blue}{Image Reconstruction: Results and Evaluation}}

In this section, \textcolor{blue}{we discuss the results of phase retrieval and image reconstruction by our cGAN model followed by an analysis of these results using moments. The hyperparameter values that produce the best performance in our cGAN  model are tabulated in Table~\ref{tab:hyperparameters}}.

\iffalse
\st{In this section, we begin by discussing ......in which multiple sources are trained simultaneously. The best performance for image reconstruction has been observed with a learning rate of \(2 \cdot 10^{-4}\), a kernel size of 5×5, and equal noise percentages (alpha/beta = 1) applied to both the original and generated images. A batch size of 1 is used, and equal training is provided to both the Discriminator and the Generator.}
\fi

\subsection{Predicted Image from the Trained cGAN}
Fig.~\ref{fig:GAN} demonstrates the success of the trained cGAN model \st{in training a model} in reconstructing the images of  fast-rotating stars using their Intensity Interferometry (II) observation . The GAN was trained on the training datasets for 100,000 steps and subsequently tested on various validation datasets to produce predicted images of the stars. The training was performed on a CPU using two nodes, each with 96 threads, and required approximately 20 hours. In Fig.~\ref{fig:GAN}, four combined images illustrate the GAN's performance in reconstructing the stars' shape, size, and brightness distribution using II.

\textcolor{blue}{The four panels from left to right in each row of Fig.~\ref{fig:GAN} represent the following:}
\begin{itemize}
\iffalse
\st{\item{The left panel shows the signals collected from six baselines, which serve as the input for the Generator during training.}}
\fi
\item{\textcolor{blue}{The left panel represents the sparse II pattern obtained by the simulated observation using the four IACTs illustrated in Fig.\ref{fig:teles}. As explained earlier, this image acts as part of the input, namely its ``condition'' part. This image acts as the condition that the sky-images of the star generated by the Generator must conform to.}}

\item{The second panel from left displays the real image, or ground truth, which the Disciminator loss function uses to distinguish from the images generated by the Generator. During training, the GAN aims to keep minimum difference between these ground truth image and generated images. \textcolor{blue}{The data generated for training, validation and testing of the cGAN model is a merge of this image and its II pattern presented in the left panel. }}

\item{The third panel from left presents the reconstructed, or predicted, image produced by the trained GAN, highlighting its success in image reconstruction.}

\item{\textbf{\textcolor{magenta}{Check if correct: The right panel shows the difference between the ground truth and the predicted image in the interferometric plane. The uniform background color of the frames suggests that the GAN model picks up white noise in the interferometric plane. I also think that axis labels need to be provided.}}}
\end{itemize}

%\iffalse

The predicted images \textcolor{blue}{presented in the third column (from left) of Fig.~\ref{fig:GAN} show} encouraging results, \st{accurately} conveying visual information about the source's size, shape, and brightness distribution across its surface \textcolor{blue}{fairly accurately. This has been achieved on the basis of the input provided by II observtion} using only six baselines. \st{However,} \textcolor{blue}{Arguably,} further improvements can be achieved by increasing the number of telescopes to maximize coverage of the $(u, v)$ plane. \textcolor{blue}{A closer examination of this proposition might be able to contribute to the design and instrumentation aspects in the existing and upcoming CTAO.} \st{an ideal candidate for this approach.}

%\fi

\begin{figure*}
	\includegraphics[width=.32\linewidth]{fig/moments/mom0.png}\hfil
	\includegraphics[width=.32\linewidth]{fig/moments/mom1.png}\hfil
	\includegraphics[width=.32\linewidth]{fig/moments/mom2.png}\hfil 
\iffalse
	\caption{This set of figures shows the comparison of monopole, x-centroid, and y-centroid for ground truth and predicted images generated by trained GAN. The monopole shows \textcolor{blue}{an approximate} linear nature, which indicate \textcolor{blue}{a correlation between the parameters of the ellipsoids and the flux (monopole) moment.} \st{that as the parameters of ellipsoid will change, the flux (monopole) also change accordingly.}\textcolor{magenta}{\textbf{Unclear statement: Clustering nature around the value of 63 for both x and y centroid show the spatial center of mass of the stellar objects.}}}
\fi
\caption{This set of figures shows the comparison of monopole, x-centroid, and y-centroid for ground truth and predicted images generated by trained GAN. The monopole shows \textcolor{blue}{an approximate} linear nature, which indicate \textcolor{blue}{a correlation between the parameters of the ellipsoids and the flux (monopole) moment.}\textcolor{magenta}{\textbf{Unclear statement: Clustering nature around the value of 63 for both x and y centroid show the spatial center of mass of the stellar objects.}}}
	\label{fig:cen}
\end{figure*}

\subsection{Evaluation of GAN using Moments}
The reconstructed images are visually compelling, demonstrating the GAN's effectiveness in using II to reconstruct images. However, visual assessment alone is insufficient; statistical evaluation is necessary to validate the results. To achieve this, we employ image moments as a statistical method. Image moments capture key properties of the reconstructed objects—such as shape, size, and intensity distribution—by quantifying features like position, orientation, and brightness distribution. By comparing the moments of the GAN-generated images to those of the ground truth, we can objectively assess the consistency and accuracy of the reconstruction. This approach provides a reliable framework for evaluating reconstruction quality, as image moments can reveal subtle differences in geometric and intensity properties that might not be apparent through visual inspection alone.

The raw moment $M_{ij}$ of an image is defined as \citep{hu1962visual}
\begin{equation}
	M_{ij} = \sum_{x} \sum_{y} x^i y^j I(x, y)
	\label{eqn:Mom}
\end{equation}
where $I(x,y)$ represents the intensity at pixel $(x,y)$. The zeroth order raw moment, or monopole, represents the total intensity of an image. It is computed by summing all pixel values across the image, yielding an overall intensity measure. In this context, analyzing the monopole provides the total flux of fast-rotating stars. According to Eq.~\eqref{eqn:Mom}, the monopole of an image is calculated as:
\begin{equation}
	M_{00} = \sum_{x} \sum_{y} I(x, y).
\end{equation}
The left panel of Fig.~\ref{fig:cen} displays the monopole values for 100 reconstructed images. The plot reveals \textcolor{blue}{an approximate} linear relationship between the monopole of the ground truth (real image) on the $x$-axis and that of the predicted (reconstructed) image on the $y$-axis, consistent across sources of varying shapes and sizes. This linearity confirms that the predicted images have an overall intensity (flux) that closely matches the ground truth. However, while the monopole effectively represents the total brightness, it does not provide information about the position, shape, size, or detailed brightness distribution of the fast-rotating stars. For these aspects, higher-order moments are necessary.

The ``center of mass'' of the sky image of any stellar object, is determined by its centroid, which provides the $x$ and $y$ coordinates representing the spatial position of the image. This centroid is computed using the first-order raw moments in conjunction with the monopole (the zeroth-order moment). The formulations for the centroid along the $x$ and $y$ directions are given by:
\begin{eqnarray}
&&x_c = \frac{M_{10}}{M_{00}} = \frac{\sum_{x,y} x \cdot I(x,y)}{\sum_{x,y} I(x,y)} \nonumber \\
&&y_c = \frac{M_{01}}{M_{00}} = \frac{\sum_{x,y} y \cdot I(x,y)}{\sum_{x,y} I(x,y)}
\end{eqnarray}
Here, \(I(x,y)\) represents the intensity at pixel \((x,y)\), \(M_{00}\) is the monopole (total intensity), and \(M_{10}\) and \(M_{01}\) are the first-order raw moments along the $x$ and $y$ axes respectively. This formulation \st{accurately} captures the spatial ``center of mass'' of the stellar object in the image.

The middle and right panel of Fig.~\ref{fig:cen} compare the centroids ($x_{c}$, $y_{c}$) of 100 predicted images with their corresponding ground truths respectively. The clustering of centroids within a specific scale range across all results indicates that the reconstructed images accurately represent the spatial location of the fast-rotating star relative to the ground truth.

\begin{figure*}
  \includegraphics[width=.32\linewidth]{fig/moments/mom3.png}\hfil
  \includegraphics[width=.32\linewidth]{fig/moments/mom4.png}\hfil
  \includegraphics[width=.32\linewidth]{fig/moments/mom5.png}\hfil
  \caption{The second-order central moments from left to right ($\mu_{11}$, $\mu_{20}$, and $\mu_{02}$) provide information about the size and shape of stellar objects. These have been shown here for ground truth and predicted images generated by the trained GAN. \st{It is very clear that these moments show the linear nature.}\textcolor{blue}{The indicative approximate linear relation arguably points towards a close correlation between the predicted and the ground truth images.}}
	\label{fig:struc}
\end{figure*}
\begin{figure*}
  \includegraphics[width=.49\linewidth]{fig/moments/mom6.png}\hfil
  \includegraphics[width=.49\linewidth]{fig/moments/mom7.png}\hfil
  \includegraphics[width=.49\linewidth]{fig/moments/mom8.png}\hfil
  \includegraphics[width=.49\linewidth]{fig/moments/mom9.png}\hfil
  \caption{Shown here are all the third-order central moments ($\mu_{30}$, $\mu_{03}$, $\mu_{21}$, and $\mu_{12}$) for ground truth and predicted images generated by the trained GAN. They represent the skewness of the brightness distributions. Here $\mu_{03}$, and $\mu_{21}$ show the linear behaviour. However, $\mu_{30}$, and $\mu_{12}$ show the clustering nature, which says that skewness is preserved but not perfectly. This aim can be achieved by increasing the number of telescopes on observational plane.}
  \label{fig:moments}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Furthermore, these calculated centroids are instrumental in analyzing the shape, size, and brightness distribution of the stars using higher-order image moments. To this end, the central moment of an image is calculated according to:
\begin{equation}
	\mu_{pq} = \frac{1}{M_{00}}\sum_{x} \sum_{y} (x - x_c)^p (y - y_c)^q I(x, y).
\end{equation}
The sum of \(p\) and \(q\) defines the order of the central moment.

Fig.~\ref{fig:struc} presents the second-order central moments (\(\mu_{11}, \mu_{20}, \mu_{02}\)), which are used to study the structure of a fast-rotating star along the line of sight (as explained in the upcoming subsection). All three plots demonstrate \textcolor{blue}{an approximately} linear relationship in the second-order moments, similar to the monopole, thereby confirming the success of applying the GAN to reconstruct images with II.

The brightness distribution is characterized by the skewness of the image, which is quantified by calculating the third-order central moments (\(\mu_{30}, \mu_{03}, \mu_{21}, \mu_{12}\)). Fig.~\ref{fig:moments} presents all third-order moments for both the ground truth and the reconstructed image. The skewness along the $x$ and $y$ axes (\(\mu_{30}\) and \(\mu_{03}\)) appears acceptable, as shown in both upper panel of Fig.~\ref{fig:moments}, where a linear relationship exists between the ground truth and predicted images. However, the other higher-order moments (\(\mu_{21}\) and \(\mu_{12}\))—particularly \(\mu_{12}\), as depicted in both lower panel of Fig.~\ref{fig:moments}—do not align as well. This indicates that further improvement is possible and should be investigated.

\subsection{The reconstructed Parameters for object}
The centroids \((x_c, y_c)\) indicate only the center of the star and its spatial location in the image. In contrast, the second-order central moments determine the orientation, semi-major axis, and eccentricity relative to the source's center \citep{teague1980image}. These moment-based parameters fully describe the two-dimensional ellipse that fits the image data.

The orientation of a fast-rotating star along the line of sight is defined in terms of second-order central moments as
\begin{equation}
	\theta = \tfrac{1}{2}\arctan \left(\frac{2\mu_{11}}{\mu_{20} - \mu_{02}}\right).
	\label{eqn:orn}
\end{equation}
The semi-major and semi-minor axes of the stellar object are computed using the second-order central moments and are denoted as \(a\) and \(b\), respectively.
\begin{equation}
	\begin{aligned}
		&a = 2\sqrt{mp + \delta} \\
		&b = 2\sqrt{mp - \delta}
	\end{aligned}
	\label{eqn:semi}
\end{equation}
where,
\begin{equation}
	mp = \frac{\mu_{20} + \mu_{02}}{2}
	\label{eqn:mp}
\end{equation}
and
\begin{equation}
	\delta = \frac{\sqrt{4\mu_{11}^2 + (\mu_{20} - \mu_{02})^2}}{2}.	
	\label{eqn:delta}
\end{equation}
Using the calculated axis values, the eccentricity of the fast-rotating star is determined as:
\begin{equation}
	e = \sqrt{1 - a/b}.
	\label{eqn:eccen}
\end{equation}
Eqs.~\ref{eqn:orn}-\ref{eqn:eccen} describe the elliptical nature of the stellar object (in this case, a fast-rotating star) and provide information on its shape and size, depending on the computed values. In contrast, the brightness distribution is characterized by skewness, which is quantified using third and higher-order moments.