### 2025-11-27 Phase_II/draft/main.tex

\documentclass[linenumbers, twocolumn]{aastex701}
\usepackage{amsmath}
\usepackage{tikz}
\usepackage{xcolor}
%\usepackage[normalem]{ulem}
\usepackage{soul}
\usepackage{hyperref}
%\newcommand{\paperdoi}{10.5281/zenodo.17598807}

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyfoot{}
\fancyfoot[L]{DOI: \href{https://doi.org/10.5281/zenodo.17598807}{10.5281/zenodo.17598807}}

\usetikzlibrary{shapes, arrows.meta, positioning, calc}

\tikzset{
	block/.style = {draw, rectangle, minimum height=1.5em, minimum width=3em, align=center},
	arrow/.style = {->, thick},
}

\definecolor{BLUE}{named}{blue}

\newcommand{\mias}{$\SI{}{\micro{\rm as}}$}
\newcommand{\abs}[1]{\left|#1\right|}
\newcommand{\norm}[2]{\left\lVert#1\right\rVert}

\begin{document}

\title{Generative AI for image reconstruction in Intensity Interferometry: a first attempt}

\author{Km Nitu Rai}
\affiliation{School of Physics, Indian Institute of Science Education and Research Thiruvananthapuram, Maruthamala PO, Vithura, Thiruvananthapuram 695551, Kerala, India.}
\affiliation{Aryabhatta Research Institute of Observational Sciences, Manora Peak, Nainital 263129, India.}
\email{niturai201296@gmail.com}

\author{Yuri van der Burg} 
\affiliation{Physik-Institut, University of Zurich,	Winterthurerstrasse 190, 8057 Zurich, Switzerland.}
\email{yuri.vanderburg@uzh.ch}

\author{Soumen Basak}
\affiliation{School of Physics, Indian Institute of Science Education and Research Thiruvananthapuram, Maruthamala PO, Vithura, Thiruvananthapuram 695551, Kerala, India.}
\email{sbasak2006@gmail.com}

\author{Prasenjit Saha}
\affiliation{Physik-Institut, University of Zurich,	Winterthurerstrasse 190, 8057 Zurich, Switzerland.}
\email{psaha@physik.uzh.ch}

\author{Subrata Sarangi}
\affiliation{School of Applied Sciences, Centurion University of	Technology and Management, Odisha-752050, India.}
\affiliation{Visiting Associate, Inter-University Centre for Astronomy and Astrophysics, Post Bag 4, Ganeshkhind, Pune 411 007, Maharashtra, India.}
\email{ssarangi.itech@gmail.com}

%\maketitle
%\begin{center}
%\textbf{DOI:} \href{https://doi.org/\paperdoi}{\paperdoi}
%\end{center}

\begin{abstract}
In the last few years Intensity Interferometry (II) has made significant strides in achieving high-precision resolution of stellar objects at optical wavelengths. Despite these advancements, phase retrieval remains a major challenge due to the nature of photon correlation. This paper explores the application of a conditional Generative Adversarial Network (cGAN) to tackle the problem of image reconstruction in II. \textcolor{blue}{This method successfully reconstructs the shape, size, and brightness distribution of simulated, fast-rotating stars based on a sparsely sampled spatial power spectrum obtained from a hypothetical ground-based II facility composed of four Imaging Atmospheric Cherenkov Telescopes (IACTs).}  Although this particular example could also be addressed using parameter fitting, the results suggest that with larger arrays much more complicated systems could be reconstructed by applying machine-learning techniques to II.
\end{abstract}

\input{intro_reformatted}
\input{image-reconstruction-past-efforts}
\input{II}
\input{GAN}
\input{ParaGan}
\input{Result}
\input{conclusion}
\begin{acknowledgements}
\textbf{Acknowledgment:} One of the authors (SS) gratefully acknowledges the computing facilities and the local hospitality extended to him by the Inter-University Centre for Astronomy and Astrophysics (IUCAA), Pune, India under its Visiting Associate Programme during the preparation of this manuscript.
\end{acknowledgements}

\bibliographystyle{aasjournalv7}
\bibliography{refs}

\end{document}