### 2025-12-07 Phase_II/draft/Result.tex

\section{Image Reconstruction: Results and Evaluation}

In this section, we examine the performance of the GAN Model whose architecture and choice of hyperparameters have been discussed above. We subject the trained Model to the task of phase retrieval and image reconstruction on the Testing Dataset.

\subsection{Visual Evaluation of Images Predicted by the Model}
Fig.~\ref{fig:GAN} demonstrates the success of the trained Model in reconstructing the images of a sample of the fictitious fast-rotators drawn from the Testing Set.

The four panels from left to right in each row of Fig.~\ref{fig:GAN} represent the following:
\begin{itemize}
\item{The left panel represents the sparse II pattern obtained by the simulated observation using the four IACTs illustrated in Fig.\ref{fig:teles}. As explained earlier, this image acts as part of the input, namely its ``condition'' part. This image acts as the condition that the sky-images of the star generated by the Generator must conform to.}

\item{The second panel from left displays the real image, or ground truth target image, which the Discriminator loss function uses to distinguish from the images generated by the Generator.}

\item{The third panel from left presents the reconstructed (or predicted) image corresponding to the ground truth (second panel) and generated by the trained Model. The similarity of these two images indicates the success of the Model in its stated objective of image reconstruction.}

\item{The right panel shows the residuals between the ground truth target image and the predicted image in the interferometric plane. The small values are indicative of the success of training the Model.}
\end{itemize}

The predicted images and the residuals presented in the third and the fourth columns (from left) of Fig.~\ref{fig:GAN} show encouraging results, conveying visual information about the source's size, shape, and brightness distribution across its surface fairly accurately. This has been achieved on the basis of the input provided by II observation using only six baselines. Arguably, further improvements can be achieved by increasing the number of telescopes to enhance the coverage of the $(u, v)$ plane. A closer examination of this proposition might be able to contribute to the design and instrumentation aspects in the existing and upcoming CTAO.

\begin{figure*}
	\includegraphics[width=.32\linewidth]{fig/moments/mom0.png}\hfil
	\includegraphics[width=.32\linewidth]{fig/moments/mom1.png}\hfil
	\includegraphics[width=.32\linewidth]{fig/moments/mom2.png}\hfil

\caption{This set of figures shows comparisons of the monopole, and the coordinates of the centroids ($x_c,\, y_c$) of the pixel brightness distribution of the ground truth images and the corresponding images predicted (generated) by the trained model. The monopole moment of the images represents the overall pixel brightness of the images. The values, as expected of a trained Model, not only lie in same range, but lie close to the diagonal and show an approximate equality. The centroids of both the ground truth images and the predicted images lie within small ranges of pixel values as should be expected of an ideally trained model. Please see the relevant text for detailed comments.}
\label{fig:cen}
\end{figure*}

\subsection{Evaluation of the Model using Moments}
The reconstructed images are visually compelling, demonstrating the Model's effectiveness in using II to reconstruct images. However, visual assessment alone is insufficient; statistical evaluation of the generated images in comparison with the ground truth images is necessary to validate the results. We present here the results of our calculation and comparison of the moments of distribution pixel brightness in the target ``Ground Truth'' images and the ``Predicted'' images.

Image moments capture key properties of the reconstructed objects, such as shape, size, and intensity distribution, by quantifying features like position, orientation, and brightness distribution. By comparing the moments of the Model-generated images with those of the ground truth, we can objectively assess the consistency and accuracy of the reconstruction. This approach provides a reliable framework for evaluating reconstruction quality, as image moments can reveal subtle differences in geometric and intensity properties that might not be apparent through visual inspection alone.

The raw moment $M_{ij}$ of an image is defined as \citep{hu1962visual}
\begin{equation}
	M_{ij} = \sum_{x} \sum_{y} x^i y^j I(x, y)
	\label{eqn:Mom}
\end{equation}
where $I(x,y)$ represents the intensity at pixel $(x,y)$. The zeroth order raw moment, or monopole, represents the total intensity of an image. It is computed by summing all pixel values across the image, yielding an overall intensity measure. In this context, analyzing the monopole provides the total pixel brightness of the images of the fictitious stars. According to Eq.~\eqref{eqn:Mom}, the monopole of an image is calculated as:
\begin{equation}
	M_{00} = \sum_{x} \sum_{y} I(x, y).
\end{equation}

An ideally trained Model should predict the overall pixel brightness of the generated images to be in the same range of values, if not equal to those of the ground truth images. The left panel of Fig.~\ref{fig:cen} displays the monopole values for 100 reconstructed images. The plot reveals a broad agreement in the overall intensity of the ground truth targets and the predicted (generated) images. However, a slightly higher brightness of reconstructed images in the intermediate range of brightness is evident. This could, perhaps, be mitigated by tuning the Batch Size hyperparameter.

While the monopole effectively represents the total brightness, it does not provide information about the position, shape, size, or detailed brightness distribution of the fast-rotating stars. For these aspects, higher-order moments are necessary.

The first order raw moments normalized by the respective monopole moments of the images given by 
\begin{eqnarray}
&&x_c = \frac{M_{10}}{M_{00}} = \frac{\sum_{x,y} x \cdot I(x,y)}{\sum_{x,y} I(x,y)} \nonumber \\
&&y_c = \frac{M_{01}}{M_{00}} = \frac{\sum_{x,y} y \cdot I(x,y)}{\sum_{x,y} I(x,y)}
\end{eqnarray}
represent the centroids $(x_c, y_c)$ of the pixel brightness distribution of the images. An ideally trained GAN model should produce the centroids of the target and predicted images within a small range of pixels. A larger range would imply a low fidelity in learning the brightness distribution of ground truth images. The middle and right panels of Fig.~\ref{fig:cen} compare the centroids ($x_{c}$, $y_{c}$) of 100 predicted images with their corresponding ground truths respectively. We notice that the coordinates of the centroids of ground truth images are spread across a range of 2 pixels whereas the those of the predicted images are spread across a span of 4 pixels along $x$-axis and 7 pixels along $y$-axis. This clustering of centroids within a small scale range across all results indicates that the reconstructed images represent the spatial distribution of brightness of predicted and ground truth images fairly accurately.

\begin{figure*}
  \includegraphics[width=.32\linewidth]{fig/moments/mom3.png}\hfil
  \includegraphics[width=.32\linewidth]{fig/moments/mom4.png}\hfil
  \includegraphics[width=.32\linewidth]{fig/moments/mom5.png}\hfil
  \caption{The second order central moment of the brightness distribution of the images is analogous to the moment of inertia of a mass distribution. The panels of this figure show the comparison of the second order moments ($\mu_{11}$, $\mu_{20}$, and $\mu_{02}$) of the predicted images vs. the ground truth images from left to right respectively. Approximate equality of these moments is evident. The small scatter in the moments of the predicted images is indicative of a balanced training of the Model. }
\label{fig:struc}
\end{figure*}
\begin{figure*}
  \includegraphics[width=.49\linewidth]{fig/moments/mom6.png}\hfil
  \includegraphics[width=.49\linewidth]{fig/moments/mom7.png}\hfil
  \includegraphics[width=.49\linewidth]{fig/moments/mom8.png}\hfil
  \includegraphics[width=.49\linewidth]{fig/moments/mom9.png}\hfil
  \caption{Shown here are all the third-order central moments ($\mu_{30}$, $\mu_{03}$, $\mu_{21}$, and $\mu_{12}$) for ground truth and predicted images generated by the trained Model. They represent the skewness of the brightness distributions. The low skewness in the brightness distribution of the ground truth images is evident. One outlier among the predicted images dominates their skewness distribution along the $y$-axis as seen in the values of \(\mu_{03}\) and \(\mu_{12}\). The trained Model, however, seems to have learned the approximately uniform distribution of skewness along the $x$-axis as is evident, with a few outliers though, in the values of \(\mu_{30}\) and \(\mu_{21}\).}
\label{fig:moments}
\end{figure*}

Furthermore, these calculated centroids are instrumental in analyzing the shape, size, and brightness distribution of the stars using higher-order image moments. To this end, the central moment of an image is calculated according to:
\begin{equation}
	\mu_{pq} = \frac{1}{M_{00}}\sum_{x} \sum_{y} (x - x_c)^p (y - y_c)^q I(x, y).
\end{equation}
The sum of \(p\) and \(q\) defines the order of the central moment.

The second order central moment of the brightness distribution of the images is analogous to the moment of inertia of a mass distribution. Fig.~\ref{fig:struc} presents the comparison of second-order central moments (\(\mu_{11}, \mu_{20}, \mu_{02}\)), which are used to study the structure of a fast-rotating star along the line of sight (as explained in the upcoming subsection). All three plots demonstrate an approximate equality among these moments, similar to the monopole, thereby confirming the success of applying the Model to reconstruct images with II. The small scatter in the moments of the predicted images indicates robust learning of the Model without over-fitting.

The third-order central moments (\(\mu_{30}, \mu_{03}, \mu_{21}, \mu_{12}\)) of the images quantify the degree and direction of asymmetry of the brightness distribution around their respective centroids. Fig.~\ref{fig:moments} presents a comparison of all third-order moments for both the ground truth and the reconstructed image.    All the four panels show a low skewness in the brightness distribution of the ground truth images. The skewness of the predicted images along the $y$-axis is dominated by one outlier as seen in the values of \(\mu_{03}\) and \(\mu_{12}\). However, the approximately uniform distribution of skewness of the predicted images along the $x$-axis, with a few outliers though, as seen in the values of \(\mu_{30}\) and \(\mu_{21}\) suggests that this feature in the ground truth images has been picked up fairly well during the training of the Model.

\subsection{The reconstructed Parameters for object}
The centroids \((x_c, y_c)\) indicate only the center of the star and its spatial location in the image. In contrast, the second-order central moments determine the orientation, semi-major axis, and eccentricity relative to the source's center \citep{teague1980image}. These moment-based parameters fully describe the two-dimensional ellipse that fits the image data.

The orientation of a fast-rotating star along the line of sight is defined in terms of second-order central moments as
\begin{equation}
	\theta = \tfrac{1}{2}\arctan \left(\frac{2\mu_{11}}{\mu_{20} - \mu_{02}}\right).
	\label{eqn:orn}
\end{equation}
The semi-major and semi-minor axes of the stellar object are computed using the second-order central moments and are denoted as \(a\) and \(b\), respectively.
\begin{equation}
	\begin{aligned}
		&a = 2\sqrt{mp + \delta} \\
		&b = 2\sqrt{mp - \delta}
	\end{aligned}
	\label{eqn:semi}
\end{equation}
where,
\begin{equation}
	mp = \frac{\mu_{20} + \mu_{02}}{2}
	\label{eqn:mp}
\end{equation}
and
\begin{equation}
	\delta = \frac{\sqrt{4\mu_{11}^2 + (\mu_{20} - \mu_{02})^2}}{2}.	
	\label{eqn:delta}
\end{equation}
Using the calculated axis values, the eccentricity of the fast-rotating star is determined as:
\begin{equation}
	e = \sqrt{1 - a/b}.
	\label{eqn:eccen}
\end{equation}
Eqs.~\ref{eqn:orn}-\ref{eqn:eccen} describe the elliptical nature of the stellar object (in this case, a fast-rotating star) and provide information on its shape and size, depending on the computed values. In contrast, the brightness distribution is characterized by skewness, which is quantified using third and higher-order moments.